---
title: "4550HW1"
author: "Xinyi Liu"
date: "February 12, 2018"
output: html_document
---

Exercise 2
1.
```{r,echo=T}
load("E:/STSCI4550/y.RData")
ex2data=y
```

2. 
part a
```{r,echo=T}
ex2ts <- ts(ex2data, frequency = 12, start = c(2008,1))
```
part b
```{r,echo=T}
ex2plot=ts.plot(ex2ts, main = "Frequency = 12")
ex2plot
```
From the ts plot above, there is no clear trend indicated.

part c
```{r,echo=T}
ex2sea <- ts(ex2data, frequency = 4, start = c(2008,1))

seaplot=ts.plot(ex2sea, main = "Frequency = 4")
seaplot
monthplot(ex2ts)
```
There is no obvious seasonality observed from the plot. Seasonal data are randomly distributed around zero. 

3.
```{r,echo=T}
library(CADFtest)
max.lag<-round(sqrt(length(ex2ts)))
CADFtest(ex2ts, type= "drift", criterion= "BIC", max.lag.y=max.lag) 
```
Because p-value=1.244e-06,we decide to rejectthe null hypothesis, the series does not have a stochastic trend, it is stable & stationary.

4. 
part a
```{r,echo=T}
acf(ex2ts) #MA(1)
```
From the correlogram, it is a MA(1) process.

part b
```{r,echo=T}
fitma <- arima(ex2ts, order = c(0,0,1))
fitma
abs(fitma$coef/sqrt(diag(fitma$var.coef)))
```
From results above, the coefficient for the fitted model is 0.4721 and the intercept is -0.3860, since 6.366928 and  2.694825 are larger than 1.96, coefficients are significant. 

part c
```{r,echo=T}
ts.plot(fitma$residuals)
acf(fitma$residuals)
Box.test(fitma$residuals, lag = max.lag, type = "Ljung-Box")

```
p-value=0.9441, not significant, so we do not reject null hypothesis, the model is validated, the residual series is a white noise.

5.
part a
```{r,echo=T}
pacf(ex2data) #AR(1)

```
Based on the correlogram,it is a AR(1) process

part b
```{r,echo=T}
fitar <- arima(ex2ts, order = c(1,0,0))
fitar
abs(fitar$coef/sqrt(diag(fitar$var.coef)))
```
The coefficient of fitted model is 0.4722 , the intercept is -0.0799, both coefficients are significant since 5.911237 and 2.086681 are large enough.

part c
```{r,echo=T}
ts.plot(fitar$residuals)
acf(fitar$residuals)
Box.test(fitar$residuals, lag = max.lag, type = "Ljung-Box")
```
Since p-value=0.9588, not significant, we do not reject null hypothesis, the model is validated and the residual series is a white noise.

6.
part a
```{r,echo=T}
AIC(fitma)
AIC(fitar)

```
From the result, we see AIC for MA model > AIC for AR model, thus we prefer model AR(1).

part b
```{r,echo=T}
BIC(fitma)
BIC(fitar)
```
From the result, we see BIC for MA model> BIC for AR model, thus we prefer model AR(1).


Exercise 3

1.
```{r,echo=T}
ex3 <- read.table(file = file.choose(), header=TRUE)
attach(ex3)
```

2.
part a
```{r,echo=T}
ex3ts <- ts(log(PUNEW), frequency = 4, start = c(1989,1))
```
part b
```{r,echo=T}
ex3plot=ts.plot(ex3ts, main = "Frequency = 4")
ex3plot
```
The time series has a clear trend from the plot.

part c
```{r,echo=T}
max.lag1<-round(sqrt(length(ex3ts)))
CADFtest(ex3ts, type= "trend", criterion= "BIC", max.lag.y=max.lag1)
```
Since p-value=0.4779, we failed to reject null hypothesis, the series is not stationary, it has a stochastic trend.

part d
```{r,echo=T}
dlogts <- diff(ex3ts)
ts.plot(dlogts)

```

part e
```{r,echo=T}
max.lag2<-round(sqrt(length(dlogts))) 
CADFtest(dlogts, type= "drift", criterion= "BIC", max.lag.y=max.lag2) 

```
Since p-value=3.826e-06, we reject null hypothesis, there is no unit root, series is stationary. 

3. 
part a
```{r,echo=T}
acf(dlogts) #MA(3)
```
From the correlogram, we can see that it is a MA(3) process.

part b
```{r,echo=T}
fitma1 <- arima(dlogts, order = c(0,0,3))
fitma1
abs(fitma1$coef/sqrt(diag(fitma1$var.coef)))
```
The coefficients of fitted model are 0.3696, 0.2031 and 0.4089, the intercept is 0.0072.
The coefficient for ma2 isn't significant, ma1 and ma3 are significant.

part c
```{r,echo=T}
ts.plot(fitma1$residuals)
acf(fitma1$residuals)
Box.test(fitma1$residuals, lag = max.lag2, type = "Ljung-Box")

```
Since p-value=0.9969, not significant, we do not reject null hypothesis, the model is validated, the residual series is a white noise.

4.
part a
```{r,echo=T}
pacf(dlogts) #AR(3)

```
From the correlogram above, we see that it is a AR(3) process.

part b
```{r,echo=T}
fitar1 <- arima(dlogts, order = c(3,0,0))
fitar1
abs(fitar1$coef/sqrt(diag(fitar1$var.coef)))
```
The coefficients are 0.3079,0.0530 and 0.2983 respectively,and the coefficient for ar2 isn't significant, ar1 and ar3 are significant.

part c
```{r,echo=T}
ts.plot(fitar1$residuals)
acf(fitar1$residuals)
Box.test(fitar1$residuals, lag = max.lag2, type = "Ljung-Box")
```
Since p-value=0.8696,not significant,  we failed to reject null hypothesis, the model is validated, the residual series is a white noise.

5.
part a
```{r,echo=T}
AIC(fitma1)
AIC(fitar1)

```
Since AIC for MA model<AIC for AR model, we prefer MA(1) model.

part b
```{r,echo=T}
BIC(fitma1)
BIC(fitar1)
```
Since BIC for MA model<BIC for AR model, we prefer MA(1) model.